Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''

### 2023-02-20 additional_files/main_text.tex
Integrate suggestions by JA Sibbesen into the main figure captions

\begin{figure}
	\contcaption{Results of SV- and \textit{k}-mer-based GWAS at the
	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots of
	the analyses based on Paragraph (SV) genotype calls and
	\textit{k}-mers.  Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{W1} locus. (b) Zoomed-in Manhattan plots of the analyses based
	on Paragraph genotypes and \textit{k}-mers at the \textit{W1} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.13g072100 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 5\% (Paragraph) and
	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
	Rectangles represent exons, colored rectangles represent coding
	sequences, and arrows indicate the direction of transcription. The red
	rectangle highlights the region shown in panel (d). (d) Haplotypes and
	significant \textit{k}-mers at the \textit{W1} locus. Nucleotides are
	colored according to the \textit{p}-value of the most significant
	\textit{k}-mer overlapping them. Vertical lines between haplotypes
	indicate sequence differences, whereas dashes indicate gaps in the
	alignment. (e) Contingency table of the phenotypes and haplotypes found
	in the population at the \textit{W1} locus. The haplotypes are the ones
	represented in panel (d).}
\end{figure}

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots
	of the analyses based on Platypus (SNP/indel) genotype calls and
	\textit{k}-mers.  Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Td} locus. (b) Zoomed-in Manhattan plots of the analyses based
	on Platypus genotypes and \textit{k}-mers at the \textit{Td} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.03g258700 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of Platypus, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrow indicates the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Haplotypes and significant \textit{k}-mers at
	the \textit{Td} locus. Nucleotides are colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{Td} locus. The haplotypes are the ones represented in panel
	(d).}
\end{figure}

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots
	of the analyses based on Platypus (SNP/indel) genotype calls and
	\textit{k}-mers.  Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{G} locus. (b) Zoomed-in Manhattan plots of the analyses based
	on Platypus genotypes and \textit{k}-mers at the \textit{G} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.01g198500 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of Platypus, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Haplotypes and significant \textit{k}-mers at
	the \textit{G} locus. Nucleotides are colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{G} locus. The haplotypes are the ones represented in panel (d).}
\end{figure}

@@

% Figure 4 with the results on pubescence density at locus Ps
\begin{figure}
	\centering
	\includegraphics[width = 6in]{pubescence_density_Ps_main_figure.png}
	\caption[Results of \textit{k}-mer-based GWAS at the
	\textit{Ps} locus for pubescence density.]{Results of \textit{k}-mer-based GWAS at the
	\textit{Ps} locus for pubescence density.
	(a) Genome-wide Manhattan plot of significant \textit{k}-mers. The
	horizontal dotted line indicates the significance threshold whereas the
	vertical dotted line indicates the location of the \textit{Ps} locus.
	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus.  The horizontal
	dotted line represents the significance threshold whereas the
	vertical dotted line indicates the location of the most significant
	\textit{k}-mer. The shaded gray area shows the extent of the causal CNV.
	(c) Estimated copy number (sequencing depth at position divided by
	accession average sequencing depth) of accessions with normal and
	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
	the average copy number across 73 (semi-sparse pubescence) and 180
	accessions (normal pubescence) computed using a 100-bp sliding window.
	Shaded areas outline the first and third quartiles. The start and end
	of the CNV region can be clearly inferred from the divergence between
	the copy number of accessions with contrasting pubescence.
	(d) Histogram of the mapping location of 1,169 reads containing the
	most significant \textit{k}-mer in accessions with semi-sparse pubescence.
	Reads containing this \textit{k}-mer map to either end of the CNV
	region, suggesting that they are associated with this variant. Panels
	b, c and d share the same x-axis.}
	\label{pubescence-density-main-figure}
\end{figure}

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots
	of the analyses based on Platypus (SNP/indel) genotype calls and
	\textit{k}-mers. Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots of the analyses based
	on Platypus genotypes and \textit{k}-mers at the \textit{Pa1} locus.
	Horizontal dotted lines again indicate the significance threshold
	whereas the vertical dotted lines indicate the location of the
	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of Platypus, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription.
	(d) Zoomed-in Manhattan plots at the location of the Glyma.12g213900 gene.
	Data is represented on the same x-scale in these plots as the gene model
	in panel (c). Vertical lines indicate the location of two non-synonymous
	SNPs associated with appressed pubescence, which may represent causal variants
	at this locus.}
\end{figure}

@@

% Figure 6 with the results novel loci for stem termination type
\begin{figure}
	\centering
	\includegraphics[width = 6in]{stem_termination_sn_main_figure.png}
	\caption[Detection of novel signals for stem termination type
	using a \textit{k}-mer GWAS analysis contrasting accessions with
	indeterminate and semi-determinate stems]{Detection of novel signals for stem termination type
	using a \textit{k}-mer GWAS analysis contrasting accessions with
	indeterminate and semi-determinate stems.
	(a) Genome-wide Manhattan plot of significant \textit{k}-mers.
	The horizontal dotted line indicates the significance threshold
	whereas vertical dotted lines indicate the location of the \textit{Dt1}
	classical locus as well as three newly documented loci (\textit{stGm11},
	\textit{stGm16}, and \textit{stGm18}).
	(b) Pairwise LD among the 162 significant \textit{k}-mers identified.
	\textit{k}-mers are sorted along the y-axis according to their
	putative position along the reference assembly version 4 of Williams82,
	as identified by ``Gm'' chromosome labels. \textit{k}-mers are
	represented in the same order along the x- and y-axis. The colored
	rectangles drawn below the x-axis represent the
	-log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.}
	\label{stem-termination-main-figure}
\end{figure}

>>>>

\begin{figure}
	\contcaption{Results of SV- and \textit{k}-mer-based GWAS at the
	\textit{W1} locus for flower color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{W1} locus. (b) Zoomed-in Manhattan plots at the \textit{W1} locus.
	The vertical dotted lines indicate the location of the
	Glyma.13g072100 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 5\% (SVs) and
	1\% (\textit{k}-mers) associations.  (c) Gene model of Glyma.13g072100.
	Rectangles represent exons, colored rectangles represent coding
	sequences, and arrows indicate the direction of transcription. The red
	rectangle highlights the region shown in panel (d). (d) Pairwise
	alignment of the two haplotype sequences in the population at the
	\textit{W1} locus. Nucleotides are
	colored according to the \textit{p}-value of the most significant
	\textit{k}-mer overlapping them. Vertical dotted lines between haplotypes
	indicate sequence differences, whereas dashes indicate gaps in the
	alignment. (e) Contingency table of the phenotypes and haplotypes found
	in the population at the \textit{W1} locus.}
\end{figure}

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Td} locus for pubescence color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Td} locus. (b) Zoomed-in Manhattan plots at the \textit{Td} locus.
	The vertical dotted lines indicate the location of the
	Glyma.03g258700 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNP/indels, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.03g258700. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrow indicates the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Pairwise alignment of the two haplotype
	sequences in the population at the \textit{Td} locus. Nucleotides are
	colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical dotted line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{Td} locus.}
\end{figure}

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{G} locus for seed coat color. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{G} locus. (b) Zoomed-in Manhattan plots at the \textit{G} locus.
	The vertical dotted lines indicate the location of the
	Glyma.01g198500 gene associated with this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNP/indels, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.01g198500. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription. The red rectangle highlights the region
	shown in panel (d). (d) Pairwise alignment of the two haplotype
	sequences in the population at the \textit{G} locus. Nucleotides are
	colored according to the
	\textit{p}-value of the most significant \textit{k}-mer overlapping
	them. The vertical dotted line between haplotypes indicates the location of
	the causal SNP at this locus. (e) Contingency table of
	the phenotypes and haplotypes found in the population at the
	\textit{G} locus.}
\end{figure}

@@

% Figure 4 with the results on pubescence density at locus Ps
\begin{figure}
	\centering
	\includegraphics[width = 6in]{pubescence_density_Ps_main_figure.png}
	\caption[Results of \textit{k}-mer-based GWAS at the
	\textit{Ps} locus for pubescence density.]{Results of \textit{k}-mer-based GWAS at the
	\textit{Ps} locus for pubescence density.
	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers. The
	horizontal dotted line indicates the significance threshold whereas the
	vertical dotted line indicates the location of the \textit{Ps} locus.
	(b) Zoomed-in Manhattan plot at the \textit{Ps} locus. The
	vertical dotted line indicates the location of the most significant
	\textit{k}-mer. The shaded gray area shows the location of the causal CNV.
	(c) Estimated copy number (sequencing depth at position divided by
	accession average sequencing depth) of accessions with normal and
	semi-sparse pubescence at the \textit{Ps} locus. Solid lines represent
	the median copy number across 73 (semi-sparse pubescence) and 180
	accessions (normal pubescence) computed using a 100-bp sliding window.
	Shaded areas outline the first and third quartiles. The start and end
	of the CNV region can be clearly inferred from the divergence between
	the copy number of accessions with contrasting pubescence.
	(d) Histogram showing the location of 1,169 mapped reads containing the
	most significant \textit{k}-mer in accessions with semi-sparse pubescence.
	Reads containing this \textit{k}-mer map to either end of the CNV
	region, suggesting that they are associated with this variant. Panels
	(b), (c) and (d) share the same x-axis.}
	\label{pubescence-density-main-figure}
\end{figure}

@@

\begin{figure}
	\contcaption{Results of SNP/indel- and \textit{k}-mer-based GWAS at the
	\textit{Pa1} locus for pubescence form. (a) Genome-wide Manhattan plots.
	Horizontal dotted lines indicate the significance
	threshold whereas vertical dotted lines indicate the location of the
	\textit{Pa1} and \textit{Pa2} loci. (b) Zoomed-in Manhattan plots at the \textit{Pa1} locus.
	The vertical dotted lines indicate the location of the
	Glyma.12g213900 gene suggested as a candidate for this locus. The shaded gray
	rectangles show the regions delimited by the top 1\% associations.
	In the case of SNP/indels, blue points denote markers used in the
	original analysis, whereas red points denote points that had originally
	been pruned but whose \emph{p}-values were computed after signal
	discovery.
	(c) Gene model of Glyma.12g213900. Rectangles represent exons, colored
	rectangles represent coding sequences, and the arrows indicate the
	direction of transcription.
	(d) Manhattan plots at the location of the Glyma.12g213900 gene.
	Vertical lines indicate the location of two non-synonymous
	SNPs associated with appressed pubescence, which may represent causal variants
	at this locus. Panels (c) and (d) share the same axis.}
\end{figure}

@@

% Figure 6 with the results novel loci for stem termination type
\begin{figure}
	\centering
	\includegraphics[width = 6in]{stem_termination_sn_main_figure.png}
	\caption[Detection of novel signals for stem termination type
	using a \textit{k}-mer GWAS analysis contrasting accessions with
	indeterminate and semi-determinate stems]{Detection of novel signals for stem termination type
	using a \textit{k}-mer-based GWAS analysis contrasting accessions with
	indeterminate and semi-determinate stems.
	(a) Genome-wide Manhattan plot showing the location of significant \textit{k}-mers.
	The horizontal dotted line indicates the significance threshold
	whereas vertical dotted lines indicate the location of the \textit{Dt1}
	classical locus as well as three new loci of interest (\textit{stGm11},
	\textit{stGm16}, and \textit{stGm18}).
	(b) Pairwise LD between the 162 significant \textit{k}-mers identified.
	\textit{k}-mers are sorted along the axes according to their
	putative position on the reference assembly version 4 of Williams82.
	\textit{k}-mers are
	represented in the same order along the x- and y-axis. The blue
	rectangles below the x-axis represent the
	-log\textsubscript{10}(\textit{p}) of each \textit{k}-mer.}
	\label{stem-termination-main-figure}
\end{figure}
